---
title: "Implementation of SREWMA Control Chart"
author: "Jorge Merlo"
date: "6/5/2021"
output: 
  md_document: 
  variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Synopsis

In any industry, quality of a process is determined by their capacity to generate products/services that met the requirements established by the consumers. To meet these specifications multivariate statistical process control (MSPC), through control charts, evaluates a multivariate process quality by monitoring its underlying distribution in real-time, the purpose of this is to detect as soon as possible any potential mean or scale shift attributed to special causes.

Although the traditional Hotelling's $T^2$ is the most common control chart in MSPC, it is built under the assumption that process follows a multivariate normal distribution. Nevertheless, it is well known that in practice this assumption is rarely fulfilled because of the process often following an unknown distribution.

Therefore, multivariate nonparametric approaches such as the Signed Rank Exponentially Weighted Average (SREWMA) control chart **(cita)** can be considered as an efficient alternative that allows the monitoring of processes for which no known distribution is assumed. In this document we discuss the implementation of SREWMA Control Chart to an important process monitoring problem in a semiconductor manufacturing industry. The dataset is available from the UC Irvine Machine Learning Repository (http://archive.ics.uci.edu/ml/
machine-learning-databases/secom/secom.names).

# Data Preprocessing

In order to make a simple reproduction of the following analysis, instead of working with data that comes from a local .csv file we directly download the data from the url and load it into the environment:

```{r pressure, echo=FALSE}
download.file("http://archive.ics.uci.edu/ml/machine-learning-databases/secom/secom.data", destfile = "secom.data") # explanatory variables

download.file("http://archive.ics.uci.edu/ml/machine-learning-databases/secom/secom_labels.data", destfile  = "secom_labels.data") # output

secom = read.table("secom.data")
secom_lab = read.table("secom_labels.data") 
secom = cbind(secom,secom_lab) # concatenation of variables
```

There are originally 1567 observations and 591 variables, but the dataset contains many missing values. Therefore, we process and clean the data before using it in our illustration. First, we remove the variables with constant value. We also remove the variables having 5 or more missing values:

```{r}
col.na = sapply(secom, function(x) sum(is.na(x))) # counting the number of NA's in each variable
sum(col.na>=5) # detecting which columns contain more than 5 missing values
secom = secom[, which(col.na<5)] # filtering data based on the previous condition
```

