---
title: "A Real Data Application of SREWMA Control Chart"
author: "Jorge Merlo"
date: "6/5/2021"
output:
  github_document:
  html_preview: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Statement

In any industry, quality of a process is determined by their capacity to generate products/services that met the requirements established by the consumers. To meet these specifications multivariate statistical process control (MSPC), through control charts, evaluates a multivariate process quality by monitoring its underlying distribution in real-time, the purpose of this is to detect as soon as possible any potential mean or scale shift attributed to special causes.

Although the traditional Hotelling's $T^2$ is the most common control chart in MSPC, it is built under the assumption that process follows a multivariate normal distribution. Nevertheless, it is well known that in practice this assumption is rarely fulfilled because of the process often following an unknown distribution. 

Therefore, multivariate nonparametric approaches such as the Signed Rank Exponentially Weighted Average (SREWMA) control chart Zou et al. (2012) can be considered as an efficient alternative, since allows us to monitor the mean of a multivariate processes for which no known distribution is assumed. 

In this document we reproduce the SREWMA control chart, this has been one of the most referenced works in nonparametric MSPC since is a pioneering proposal that can start monitoring with a little amount of historical observations. A ready-to-use function is provided and if necessary, it can be easily adapted to the needs of the practitioner. We discuss its implementation to a real dataset from a white wine production process. The data set contains a total of 4898 observations, and is publicly available in the UC Irvine Machine Learning Repository (http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality).

## Exploratory Analysis

In order to make a simple reproduction of the following analysis, instead of working with data that comes from a local .csv file we directly download the data from the url and load it into the environment:

```{r}
download.file("http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv", destfile = "wq_data.csv") # explanatory variables

D <- read.csv("wq_data.csv", sep = ";")

head(D)
```

The data were recorded by a computerized system, which automatically manages the process of wine sample testing from producer requests to laboratory and sensory analysis. For each observation, there are eleven continuous measurements (based on physicochemical tests) including fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulphates, and alcohol (denoted by $x_1, x_2, . . . , x_{11}$, respectively). A categorical variable, quality, indicating the wine quality between 0 (very bad) and 10 (excellent) is also provided based on sensory analysis. The goal of this data analysis is mainly to model and monitor wine quality based on physicochemical tests, more detailed discussion about this dataset is given by Cortez et al. (2009) and the references therein.

Under the SPC context of sequentially monitoring the wine production process, we assume that the in control and out of control observations and are those whose standard quality level is 7 (LV7; as also suggested by Cortez et al. (2009)) and 6 respectively. The sample correlation matrix of this data (shown below) contains several large entries, which demonstrates that the variables have considerable interrelationships and consequently a multivariate control chart is likely to be more appropriate than a univariate control chart.  

```{r}
library("corrplot")
dic = D[which(D[,ncol(D)] == 7),(1:(ncol(D)-1))] # In control data
cor.dic = cor(dic)
corrplot(cor.dic)
```

Plots below show the scatter plots of three randomly selected variables. The joint distribution of each pair of variables are far from bivariate normal. The normal Q-Q plots for these three distributions are also shown, which clearly indicate that the marginals are not normal either:

```{r}
set.seed(123)

par(mfrow = c(2,3))
w <- combn(3,2)
nam <- colnames(dic)

for (i in 1:6) {
  
  if (i <= 3){
    plot(dic[,w[1,i]], dic[,w[2,i]], pch = 16, xlab = nam[w[1,i]], 
         ylab = nam[w[2,i]], cex.lab = 1.5, cex.axis = 1.5)
  } else {
    qqnorm(dic[,(i-3)], ylab = sprintf("%s Sample Quantiles", 
                                             ylab = nam[w[2,(i-3)]]), 
           cex.lab = 1.5, cex.axis = 1.5)
    qqline(dic[,(i-3)], col = "red")
  }
  
}
```

We also run the Shapiro-Wilk goodness-of-fit test for normality and the Mardia’s multivariate normality test:

```{r}
library(MVN)
mvn(dic)
```

All these tests together with the plots shown above suggest that the multivariate normality assumption is invalid and thus we could expect the nonparametric chart to be more robust and powerful than normal-based approaches for this dataset.

## SREWMA Control Chart Implementation

In order to implement the SREWMA control chart lets assume that we have only $m  = 20$ historical observations from LV7 and initially monitored 30 observations from LV7 and then obtained the LV6 observations sequentially. The location parameter is of the greatest interest and thus we construct the SREWMA control charts to monitor the wine quality. We set the control limit $h = 22.918$ (solid horizontal line) to obtain a false alarm every 500 in control monitoring points (see Qiu (2014) for a detailed discussion on the performance of a control chart), the values of $\lambda$ are chosen to be 0.025 for the SREWMA to ensure their IC robustness to this non-normal data, Zou et al. (2012) provides tables with $h$ and $\lambda$ parameters corresponding to other specific scenarios.

```{r}
library(SpatialNP)
library(depth.plot)

set.seed(123)

m <- 20 # number of historical observations considered in control
nic <- 30 # number of observations to be monitored that are considered in control
noc <- 50 # number of observations to be monitored that are considered out of control
lambda <- 0.025 # lambda value

p <- ncol(dic) # number of dimensions

r.ic = sample(1:nrow(dic), (m+nic))

xr = dic[r.ic[1:m], ] # randomly selection for historical observations
xr <- as.matrix(xr)

xic = dic[r.ic[(m+1):(m+nic)], ] # randomly selection for historical observations
xoc = D[which(D[,ncol(D)] == 6),(1:(ncol(D)-1))] # in control observations
xoc = xoc[sample(1:nrow(xoc), noc),] # out of control observations

xi <- rbind(xic, xoc) # concatenation of monitoring points
xi <- as.matrix(xi)

ni <- nic + noc

S0 <- cov(xr)

M0 <- chol(solve(S0))
xrt <- as.matrix(xr)%*%M0

sr_m0 <- SpatialNP::spatial.rank(xrt, shape = F) # estimation of spatial ranks 
                                                 # for historical observations

# Parameters to estimate the monitoring statistics
sr_m0_2 <- sr_m0^2
RE0 <- apply(sr_m0_2, 1, sum)
RE0 <- sum(RE0)

sr_t <- c()

REt <- matrix(0, ni, p)

vt <- matrix(0, ni+1, p)

Qt <- c()

for (t in 1:ni) {
  S = cov(xr)
  M = chol(solve(S))
  xut = rbind(xr,xi[t,])%*%M
  
  sr_t = depth.plot::spatial.rank(xut[nrow(xut), ], xut[1:(nrow(xut)-1),])
    
  REt[t,] = sr_t^2
  REt = rbind(REt, rep(0,p))
  sREt = apply(REt, 1, sum)
  sREt = sum(sREt)
    
  eps_t = (RE0 + sREt)/nrow(xut)
  
  xr = rbind(xr, xi[t,])
    
  vt[(t+1),] = ((1-lambda)*vt[t,])+((lambda)*sr_t)
    
  vt = rbind(vt, rep(0, p))
    
  Qt = c(Qt, (((2-lambda)*p)/(lambda*eps_t))*sum(vt[(t+1),]^2))
  
  # print(t)
  
}

par(cex.axis = 1.5)
plot(1:ni, Qt, pch = 20, type = 'o', main = "SREWMA Control Chart for WQD Data", xlab = "Time",
     cex.lab = 1.5, cex.main = 2)
abline(h = 22.918)
abline(v = 30, lty = 2)
```

## Conclusions
From the plot, it can be seen that the SREWMA chart exceeds its control
limit from around the 34th observation (the 14th out of control LV6 observation) onwards. This excursion suggests that a marked step-change has occurred as we intended, no false alarm is presented and a quick alarm is issued as process real status is out of control (vertical dotted line). This alarm is issued once the monitoring statistic (black dot) is greater than a certain control limit (solid line). A deeper theoretical analysis and performance assessment of SREWMA control chart can be consulted in Zou et al. (2012).

## References
P. Cortez, A. Cerdeira, F. Almeida, T. Matos, and J. Reis, Modeling wine preferences by data mining from physicochemical properties, Decision Support Syst 47 (2009), 547–553.

Qiu, P. (2014). Introduction to statistical process control. CRC Press.

Zou, C., Wang, Z. & Tsung, F. (2012). A spatial rank-based multivariate ewma control chart.NavalResearch Logistics (NRL),59(2), 91–110

