---
title: "Implementation of SREWMA Control Chart"
author: "Jorge Merlo"
date: "6/5/2021"
output: 
  md_document: 
  variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

In any industry, quality of a process is determined by their capacity to generate products/services that met the requirements established by the consumers. To meet these specifications multivariate statistical process control (MSPC), through control charts, evaluates a multivariate process quality by monitoring its underlying distribution in real-time, the purpose of this is to detect as soon as possible any potential mean or scale shift attributed to special causes.

Although the traditional Hotelling's $T^2$ is the most common control chart in MSPC, it is built under the assumption that process follows a multivariate normal distribution. Nevertheless, it is well known that in practice this assumption is rarely fulfilled because of the process often following an unknown distribution.

Therefore multivariate nonparametric approaches such as the Signed Rank Exponentially Weighted Average (SREWMA) control chart **(cita)** can be considered as an efficient alternative that allows the monitoring of processes for which no known distribution is assumed. In this document we discuss the implementation of SREWMA Control Chart.
```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
